Pragmatics account for apparent limits on theory of mind use in adults

### 2016-01-28 writing/cogsci-submission/cogsci_keysar.tex
Update abstract, add some notes

\begin{abstract}
?

\textbf{Keywords:} 
?
\end{abstract}

@@

While these results are compelling, the paradigm has been criticized from several different angles. First, \citeA{HellerGrodnerTanenhaus08_Perspective} has pointed out that in many cases, the hidden object was a better fit for the referring expression than the one in common ground (e.g. the hidden roll of tape vs. the cassette tape for ``the tape''), making the hidden object \emph{a priori} more likely to be the referent. Although relative fit has not been empirically evaluated, it would be fairer to compare two objects that fit the referring expression equally well. Second, the viewpoint asymmetry paradigm is somewhat unnatural: common ground is typically built incrementally over the course of an interaction rather than presented all at once, and it is rare for a shared display to differ in perceptual accessibility \cite{HannaTanenhausTrueswell03_CommonGroundPerspective}. Third, none of the experiments reported by Keysar and colleagues included a key comparison condition where the critical item (e.g. the largest measuring cup) was \emph{also} in common ground \cite{BrownSchmidtHanna11_IncrementalPerspectiveTaking}---perhaps people would refer to the critical item less in the privileged condition than in the full common ground condition, which would put a more positive spin on the results. %Note that none of these criticisms invalidate the paradoxical result that listeners move an occluded object, they just suggest an alternative explanation. Reference disambiguation is probabilistic: common ground and goodness of fit are weighed against one another.

@@

\begin{figure}[b!]
\begin{center}
\includegraphics[scale=.75]{images/exp2MouseTracking.pdf}
\caption{Mean percentage of time spent hovering over the distractor and target cells in the experimental and baseline conditions. Error bars are 95\% confidence intervals. \todo[inline]{TODO: Collapse mouse-tracking plots to a single figure}}
\label{fig:exp2hover}
\end{center}
\end{figure}

@@

More dramatically, we find that patterns of errors in Exp.~2 diverge significantly from a uniform distribution across items, $\chi^2(7) = 24.8, p < 0.001$. Looking more closely at these patterns, we see that the only items where errors are consistently made are those where the costly over-informative utterance remains ambiguous. For example, in item 5, both the target and distractor are hair brushes: one is round and one is flat. ``Hair brush'' is already a label at the sub-class level. If we remove the two most difficult items (the ``hair brushes'' and the ``erasers''), the total percentage of errors on experimental trials drops from 26\% to 10\% and the percentage of participants making at least one error drops from 60\% to 32\%.

When we conducted a mouse-tracking analysis identical to the one reported for Exp.~1, we again found a significant interaction between condition and cell type on hover time, $b = -0.18, t(108) = -5.2, p < 0.001$. However, the effect appears to be attenuated: in experimental trials, participants spend significantly more time hovering over the target than the distractor (see Figure \ref{fig:exp2hover}) whereas in Exp.~1, participants spent roughly the same amount of time hovering over each. Due to the relatively small sample size of Exp.~1, however, we do not have the power to show this effect statistically.

@@

In a mixed model including random intercepts for raters and items, we found a significant crossover interaction of experiment and referent on mean fitness rating, $b = -0.33, t = 32, p < 0.001$. In Exp.~1, the scripted label fit the hidden distractor better than the target, but the ordering reversed in Exp.~2: the unconstrained labels fit the target much better and the hidden distractor much worse (see Fig. \ref{fig:fitnessInteraction}).

Finally, we used a mixed effects logistic regression model to estimate the effect of distractor fit and speaker constraints on the probability of making an error on critical trials. We included random intercepts for game and item ids, and an interaction term for label fitness and label referent. We find a significant interaction of referent and fitness ratings on error rates: matchers are significantly more likely to make an error when the label fits the distractor more and the target less, $b =-1.39, z = -8.16, p < 0.001$. Additionally, participants were significantly less likely to produce errors in Exp.~2 than Exp.~1, $b = -2.10, z = -2.38, p = 0.017$.

\todo[inline]{what else should these mixed models do? Should I just use mean rating rather than every turker's ratings? Feels like using $20\times$ the degrees of freedom is wrong, somehow.  Would be nice to see if they could account for item-wise effects explicitly somehow? Like the ``bottom block'' thing which was confusable in Exp.~1 but not Exp.~2.}

@@

A related difference is our decision not to use a confederate. While a confederate is useful for reducing variation across instances of the experiment, we see several downsides. Beyond the difficulties of conducting large-scale experiments with confederates on the web, it is difficult to exactly replicate all the subtleties of the confederate's behavior that might influence the .  At a methodological level, the pair of experiments we report is a reminder that manipulations administered by a trained confederate can interact in unexpected ways with a participant's social and communicative expectations. Regardless of the experimental context, it's illuminating to see how real participants naturally interact, and

A final concern is our graphical representation of occluded cells. We told players that items behind black cells were hidden to the director, showed them side-by-side displays of the director and matcher views, and quizzed them on this detail in the short attention check that players had to pass before playing. But they may have forgotten part-way through the experiment, or failed to fully internalize it. In the in-lab version, there was an actual divider between players, so it would have been more salient. We suspect that some subset of errors in our version game aren't due to a failure of perspective-taking, but instead because of a misunderstanding of what the other person's perspective is to begin with. However, this source of errors is held constant across our two experiments, and it would only bias our estimates of errors upward.

>>>>

\begin{abstract}
People are capable of accurately reasoning about the mental states of other people using theory of mind. An influential series of language understanding experiments by Keysar and colleagues, however, showed that adults systematically failed to take the speaker's perspective into account. This has been widely interpreted as a limitation on theory of mind use. In this paper we argue that these apparent failures are in fact successes. Through a minimal pair of replications comparing scripted vs. unscripted speakers, we show that critical utterances used by Keysar and colleagues are \emph{uncooperative}: they are less informative than what a speaker would actually produce in that situation. A listener who correctly reasons about the pragmatics of the task, thereby \emph{using} theory of mind, will rationally ignore whether an object is hidden to the speaker or not and will instead rely on the precision of the utterance to disambiguate. When an uncooperative confederate flouts these expectations, this strategy produces errors; when we allow participants to naturally interact, this strategy is justified and errors are reduced.

\textbf{Keywords:} 
Theory of mind; social cognition; pragmatics
\end{abstract}

@@

While these results are compelling, the paradigm has been criticized from several different angles. 
First, \citeA{HellerGrodnerTanenhaus08_Perspective} has pointed out that in many cases, the hidden object was a better fit for the referring expression than the one in common ground (e.g. the hidden roll of tape vs. the cassette tape for ``the tape''), making the hidden object \emph{a priori} more likely to be the referent. Although relative fit was not empirically evaluated prior to our study, it would generally be fairer to compare two objects that fit the referring expression equally well. 
Second, the viewpoint asymmetry paradigm is somewhat unnatural: common ground is typically built incrementally over the course of an interaction rather than presented all at once, and it is rare for a shared display to differ in perceptual accessibility \cite{HannaTanenhausTrueswell03_CommonGroundPerspective}. 
Third, none of the experiments reported by Keysar and colleagues included a key comparison condition where the critical item (e.g. the largest measuring cup) was \emph{also} in common ground \cite{BrownSchmidtHanna11_IncrementalPerspectiveTaking}---perhaps people would refer to the critical item less in the privileged condition than in the full common ground condition, which would put a more positive spin on the results. %Note that none of these criticisms invalidate the paradoxical result that listeners move an occluded object, they just suggest an alternative explanation. Reference disambiguation is probabilistic: common ground and goodness of fit are weighed against one another.

@@

\begin{figure}[b!]
\begin{center}
\includegraphics[scale=.75]{images/exp2MouseTracking.pdf}
\caption{Mean percentage of time spent hovering over the distractor and target cells in the experimental and baseline conditions. Error bars are 95\% confidence intervals. \todo[inline]{TODO: Collapse mouse-tracking plots to a single figure showing some interaction}}
\label{fig:exp2hover}
\end{center}
\end{figure}

@@

More dramatically, we find that patterns of errors in Exp.~2 diverge significantly from a uniform distribution across items, $\chi^2(7) = 24.8, p < 0.001$. Looking more closely at these patterns, we see that the only items where errors are consistently made are those where the costly over-informative utterance remains ambiguous. For example, in item 5, both the target and distractor are hair brushes: one is round and one is flat. ``Hair brush'' is already a label at the sub-class level. If we remove the two most difficult items (the ``hair brushes'' and the ``erasers''), the total percentage of errors on experimental trials drops from 24\% to 10\% and the percentage of participants making at least one error drops from 61\% to 32\%.

\todo[inline]{rdh: I don't like these mouse-tracking analyses -- I feel like the way I'm computing hover-time is sloppy and might be skewing the results (i.e. counting number of `move' events rather than actually looking at start and end times). So I could spend some time tomorrow coding up the better way\dots}
When we conducted a mouse-tracking analysis identical to the one reported for Exp.~1, we again found a significant interaction between condition and cell type on hover time, $b = -0.18, t(108) = -5.2, p < 0.001$. However, the effect appears to be attenuated: in experimental trials, participants spend significantly more time hovering over the target than the distractor (see Figure \ref{fig:exp2hover}) whereas in Exp.~1, participants spent roughly the same amount of time hovering over each. Due to the relatively small sample size of Exp.~1, however, we do not have the power to show this effect statistically.

@@

In a mixed model including random intercepts for raters and items, we found a significant crossover interaction of experiment and referent on mean fitness rating, $b = 0.34, t = 36.4, p < 0.001$. In Exp.~1, the scripted label fit the hidden distractor better than the target, but the ordering reversed in Exp.~2: the unconstrained labels fit the target much better and the hidden distractor much worse (see Fig. \ref{fig:fitnessInteraction}).

Finally, we used a mixed effects logistic regression model to estimate the effect of distractor fit and speaker constraints on the probability of making an error on critical trials. We included random intercepts for game and item ids, and an interaction term for label fitness and label referent. We find a significant interaction of referent and fitness ratings on error rates: matchers are significantly more likely to make an error when the label fits the distractor more and the target less, $b =-1.39, z = -, p < 0.001$. Additionally, participants were significantly less likely to produce errors in Exp.~2 than Exp.~1, $b = -1.38, z = -2.41, p = 0.016$.

\todo[inline]{rdh: I still get confused about which random effects to include\dots does it make sense to include random intercepts for game id and include experiment \# as a predictor when each id only participated in one experiment?}

@@

A related difference is our decision not to use a confederate. While a confederate is useful for reducing variation across instances of the experiment, we see several downsides. Beyond the difficulties of conducting large-scale experiments with confederates on the web, it is difficult to exactly replicate all the subtleties of the confederate's behavior that might influence results. The pair of experiments we report is a reminder that manipulations administered by a trained confederate can interact in unexpected ways with a participant's social and communicative expectations. Regardless of the experimental context, it's illuminating to see how real participants naturally interact.

\todo[inline]{rdh: can take out next paragraph if hurting for space...}
A final concern is our graphical representation of occluded cells. We told players that items behind black cells were hidden to the director, showed them side-by-side displays of the director and matcher views, and quizzed them on this detail in the short attention check that players had to pass before playing. But they may have forgotten part-way through the experiment, or failed to fully internalize it. In the in-lab version, there was an actual divider between players, so it would have been more salient. We suspect that some subset of errors in our version game aren't due to a failure of perspective-taking, but instead because of a misunderstanding of what the other person's perspective is to begin with. However, this source of errors is held constant across our two experiments, and it would only bias our estimates of errors upward.
